---
title: "CalFresh/School Meals analysis"
author: "Joyce Tagal"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)
library(googlesheets4)

# Parameters
sd_link <- here::here("data/NCES School district boundaries EDGE_SCHOOLDISTRICT_TL19_SY1819/schooldistrict_sy1819_tl19.shp")

zc_link <- here::here("data/earthworks_ztca_ba/ZIPCODE.shp")

cfzip_link <- here::here("data/CalFresh/by_zip_pre_postMar15.csv")

ba_geoid <- c(
  "0606001", 
  "0606013", 
  "0606085", 
  "0606055", 
  "0606097", 
  "0606041", 
  "0606075", 
  "0606095", 
  "0606081"
  )


CA_ALBERS <- 
  "+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

#===============================================================================

```

## Download data

```{r}

cfzip <- read_csv(cfzip_link)

cfzip

```



```{r}


zip <- 
  read_sf(zc_link) 

st_crs(zip)

zip %>% 
  ggplot() +
  geom_sf()

```

```{r}

sd <- 
  read_sf(sd_link) %>% 
  filter(STATEFP == "06")

st_crs(sd)

sd %>%
  ggplot() +
  geom_sf()

```

## Check zips between Bay 

```{r}
school_meals <- 
  read_sheet(
    ss = "1fea1BSSAhbSB0HLsatDaD8wxmwsF7ezQjNvjRK-QJQI", 
    skip = 1,
    col_types = "ccccccccccccccccccccccccccccccc"
  ) %>% 
  rename(
    "district_name" = provider_addloc,
    "site_name" = provider_name
  )
school_meals
```

Check zips

```{r}
school_meals %>% 
  distinct(zip) %>% 
  pull()

zip %>% 
  distinct(ZIP_CODE_5) %>% 
  pull()

```


1. expand_grid - tidy tibble with one column = all combinations of zipcode and school districts
2. function - st_intersects. if those two boundaries intersect -> TRUE
-> filter out all zipcodes that overlap with a school district
3. st_area - give the area of each zipcode
4. st_intersection - gives the intersection area between zipcode and sd
5. st_area on the intersection - gives a percentage of the intersection area vs 
the total area of the school district.
6. Provides the ratio of children in that zipcode re: district area

```{r}
sd %>% 
  filter(str_detect(NAME, "Dublin")) 

```


```{r}

ggplot() +
  geom_sf(data = zip) +
  geom_sf(data = sd, color = "red")


```

