---
title: "CalFresh/School Meals analysis"
author: "Joyce Tagal"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(sf)

# Parameters
sd_link <- "/Users/joycetagal/GitHub/jayktee/bay-area-resources/bay_area_resources/data/NCES School district boundaries EDGE_SCHOOLDISTRICT_TL19_SY1819/schooldistrict_sy1819_tl19.shp"

zc_link <- "/Users/joycetagal/GitHub/jayktee/bay-area-resources/bay_area_resources/data/Data.gov Zipcode boundaries tl_2019_us_zcta510/tl_2019_us_zcta510.shp"

cfzip_link <- "/Users/joycetagal/GitHub/jayktee/bay-area-resources/bay_area_resources/data/CalFresh/by_zip_pre_postMar15.csv"

#===============================================================================

```

## Download data

```{r}

cfzip <- read_csv(cfzip_link)

cfzip

```

```{r}

zip <- 
  read_sf(zc_link)

zip
```

```{r}

sd <- 
  read_sf(sd_link)

sd
```



1. expand_grid - tidy tibble with one column = all combinations of zipcode and school districts
2. function - st_intersects. if those two boundaries intersect -> TRUE
-> filter out all zipcodes that overlap with a school district
3. st_area - give the area of each zipcode
4. st_intersection - gives the intersection area between zipcode and sd
5. st_area on the intersection - gives a percentage of the intersection area vs 
the total area of the school district.
6. Provides the ratio of children in that zipcode re: district area



