---
title: "School_meals_report"
author: "Chris LeBoa & Joyce Tagal"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
---
## Overview

This document analyzes the number of students eligible for Free and Reduced Price
Meals (FRPM) that will be affected by Bay Area school meal site closures over
the summer period of June-August 2020. We find that school districts closing
in early months of the summer (June and July 2020) have a higher than average
percentage (53%) of students who are FRPM-eligible, compared to an average of 
45% of FRPM across districts which are closing post-July. In June and July 2020,
the closing of school sites will affect over 250,000 FRPM eligible students.

## Data sources and Assumptions

This document uses two main data sources:

* [CA Dept of Ed](https://www.cde.ca.gov/ds/sd/sd/filessp.asp) Free and Reduced Price Meal data for school year 2019-20
* [BayAreaCommunity.org](https://www.bayareacommunity.org/#/) school meal sites info, that has been updated on a 
weekly basis from March 2020 to present.

In this analysis we make several assumptions:

* Students and their families are most able to access meal sites within their
school districts. This assumption may not hold for denser, urban areas where a 
student may access meal sites in a neighboring school district. However, given
that a not-insignificant number of school districts only allow district 
students to access the school meals, this may hold on average.
* In averaging numbers of FRPM students across a district, we take a simple
average of FRPM students across schools within the district. This assumes an
equal distribution of FRPM across schools, and a similar total enrollment 
across schools.

## Load Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(googlesheets4)
```

## Read in Data

```{r}
school_meals <- 
  read_sheet(
    ss = "1fea1BSSAhbSB0HLsatDaD8wxmwsF7ezQjNvjRK-QJQI", 
    skip = 1,
    col_types = "ccccccccccccccccccccccccccccccc"
  ) %>% 
  rename(
    "district_name" = provider_addloc,
    "site_name" = provider_name
  )
school_meals
```

## Cleaning data 

```{r}
school_meals_months <- 
  school_meals %>% 
  filter(status == "Open") %>% 
  mutate(
    end_date = 
    case_when(
      str_detect(end_date, "June") == TRUE ~ "6/30",
      str_detect(end_date, "end of school year") == TRUE ~ NA_character_,
      str_detect(end_date, "06") == TRUE ~ "6/12", 
      TRUE ~ end_date
      ),
    end_month = 
      case_when(
      str_detect(end_date, "05/|5/") == TRUE ~ "End date TBC",
      str_detect(end_date, "06/|6/") == TRUE ~ "June", 
      str_detect(end_date, "07/|7/") == TRUE ~ "July",
      str_detect(end_date, "08/|8/") == TRUE ~ "August", 
      str_detect(end_date, "Until further notice") == TRUE ~ "End date TBC",
      is.na(end_date) == TRUE ~ "End date TBC", 
      TRUE ~ end_date
      )
  ) 

school_meals_months 

```

## Basic counts of schools still open

```{r}

school_meals %>% 
  filter(!is.na(status), status != "status") %>% 
  count(status)

school_meals %>% 
  distinct(site_name)

school_meals %>% 
  filter(!is.na(status)) %>% 
  distinct(district_name) #146 school districts 


```

### Schools PEBT information details 
```{r}

school_meals %>% 
  filter(!is.na(pebt_info_yn)) %>% 
  count(pebt_info_yn)

school_meals %>% 
  filter(pebt_info_yn == "1") %>% 
  distinct(pebt_notes)


school_meals %>% 
  filter(str_detect(pebt_notes, "Flyer|flyer|flier|app")) %>% 
  distinct(pebt_notes)

school_meals %>% 
  filter(pebt_info_yn == "1") %>% 
  distinct(site_name)  #43 report 

meals_pebt <- 
  school_meals %>% 
  filter(pebt_info_yn == "1") %>% 
  group_by(district_name) %>% 
  count(links_to_cfa, links_to_CDSS, contains_original_info, pebt_notes)

# meals_pebt %>% 
#   write_csv("/Users/ChrisLeBoa/Downloads/pebt_districts.csv")

meals_pebt %>% 
  mutate(
    links_to_cfa = as.numeric(links_to_cfa),
    links_to_CDSS = as.numeric(links_to_CDSS), 
    contains_original_info = as.numeric(contains_original_info)
    ) %>% 
  ungroup() %>% 
  summarise_at(c("links_to_cfa", "links_to_CDSS", "contains_original_info"), sum, na.rm = TRUE)

```



### Schools ending summer distribution 
```{r}

#Count months by school location

school_meals_months %>% 
  summarize(total_sites = n())

school_meals_months %>% 
  summarize(total_districts = n_distinct(district_name))

school_meals_months %>% 
  count(end_month) %>% 
  rename("no_schools" = n) %>% 
  mutate(pct_schools = no_schools * 100 / sum(no_schools))

school_meals_months %>%
  group_by(end_month) %>% 
  count(district_name, name = "Number of Schools") %>% 
  count(end_month) %>% 
  rename("no_districts" = n)

school_meals %>% 
  filter(county != "county") %>% 
  distinct(county)


```

## Join with FRPM
```{r}

frlink <- here::here("data/frpm1920.xlsx")

frpm <- 
  readxl::read_xlsx(frlink, sheet = 2, skip = 1) %>% 
  rename_all(.funs = ~str_to_lower(.) %>% str_replace_all(., " ", "_") %>% str_replace_all(., "[()\\n//-]", "") %>% str_replace(., "%", "pct"))

frpm %>% 
  summarize(sum = sum(frpm_count_k12))

```

```{r}

frpm %>% 
  distinct(district_type)

frpm %>% 
  distinct(district_name)

```

### Clean district names

```{r}

school_meals_clean <- 
  school_meals_months %>% 
  filter(
    !district_name %in% 
      c("YMCA Silicon Valley", 
        "San Mateo County Library",
        "Voices College Bound",
        "Ace Charter Schools",
        "Alpha Public Schools",
        "Downtown College Prep",
        "Summit Public Schools",
        "Rocketship Public Schools"
      )
  )

school_meals_clean

```

### Checking for intersection between datasets

```{r}

intersect(
  school_meals_clean %>% 
  distinct(district_name),
  frpm %>% 
  distinct(district_name)
)

```

```{r}
districts <- school_meals_clean %>% distinct(district_name) %>% pull()

frpm_districts <- 
  frpm %>% 
  filter(district_name %in% districts) %>% 
  select(
    academic_year, 
    county_name, 
    district_name, 
    enrollment_k12, 
    free_meal_count_k12,
    percent_pct_eligible_free_k12,
    frpm_count_k12,
    percent_pct_eligible_frpm_k12
  ) %>% 
  group_by(district_name) %>% 
  summarize(
    total_enroll = sum(enrollment_k12),
    free_meal_count_k12 = sum(free_meal_count_k12),
    frpm_count_k12 = sum(frpm_count_k12)
  ) %>% 
  mutate(
    percent_pct_eligible_free_k12 = free_meal_count_k12 / total_enroll,
    percent_pct_eligible_frpm_k12 = frpm_count_k12 / total_enroll
  )

frpm_districts %>% summarise(total_enroll = sum(total_enroll))



```

## Averaging district population across school closures

Assuming an equal distribution of density across counties, we assume an average 
of students on free meals and free and reduced price meals across closing 
schools. 

```{r}

level_key <- c("June", "July", "August", "End date TBC")
school_meals_average <- 
  school_meals_clean %>%
  group_by(district_name) %>% 
  count(end_month, name = "num_schools_closing") %>% 
  group_by(district_name) %>% 
  arrange(district_name) %>% 
  add_tally(num_schools_closing, name = "sum_schools") %>% 
  left_join(
    frpm_districts,
    by = "district_name"
  ) %>% 
  mutate(
    frac_enroll = num_schools_closing / sum_schools * total_enroll,
    frac_free_meal_count_k12 = num_schools_closing / sum_schools * free_meal_count_k12,
    frac_frpm_count_k12 = num_schools_closing / sum_schools * frpm_count_k12
  ) %>% 
  group_by(end_month) %>%
  summarize(
    num_districts = n_distinct(district_name),
    num_schools_closing = sum(num_schools_closing),
    sum_schools = sum(sum_schools),
    total_frac_enroll = sum(frac_enroll),
    total_frac_free_eligible = sum(frac_free_meal_count_k12),
    total_frac_frpm_eligible = sum(frac_frpm_count_k12),
  .groups = "keep") %>% 
  mutate(
    percent_eligible_free_k12 = total_frac_free_eligible / total_frac_enroll,
    percent_eligible_frpm_k12 = total_frac_frpm_eligible / total_frac_enroll
  ) %>% 
  mutate(end_month = factor(end_month, levels = level_key))


school_meals_average

```

## Conclusion and GGPlot

The analysis shows that the 22 school districts which are closing school sites 
in July have a higher-than-average percentage of children on FRPM (~65.9%) 
compared to an average of 
`r mean(school_meals_average[school_meals_average$end_month != "July", ]$percent_eligible_frpm_k12) * 100`%. 

Overall, the percentage of students on FRPM across districts closing in June
and July is 53%, compared to 44% FRPM studentsfor districts closing in August 
and beyond. The school meal site closures directly impact districts with a higher
proportion of students eligible for free and reduced lunch.

```{r}

school_meals_average %>% 
  ggplot(aes(x = end_month, y = percent_eligible_frpm_k12)) +
  geom_col() +
  ggrepel::geom_text_repel(
    aes(label = scales::percent(percent_eligible_frpm_k12)),
    direction = "y",
    nudge_y = 0.01,
    size = 3.5
  ) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  coord_cartesian(expand = TRUE) +
  labs(
    title = "Average percentage of children eligible for FRPM in districts closing meal sites",
    y = "Average percentage children eligible for Free-Reduced Meals",
    x = "Months",
    caption = "Source: CA Dept of Ed; BayAreaCommunity.org"
  )
ggsave("percentage_districts.png", height = 5, width = 7)
  

school_meals_average %>% 
  ggplot(aes(x = end_month, y = total_frac_frpm_eligible)) +
  geom_col() +
  ggrepel::geom_text_repel(
    aes(label = scales::comma(total_frac_frpm_eligible)),
    direction = "y",
    nudge_y = 3,
    size = 3.5
  ) +
  scale_y_continuous(labels = scales::label_number_si(accuracy = 1)) +
  coord_cartesian(expand = TRUE) +
  labs(
    title = "Number of children eligible for FRPM in districts closing meal sites",
    y = "Total children eligible for Free-Reduced Meals",
    x = "Months",
    caption = "Source: CA Dept of Ed; BayAreaCommunity.org"
  )
ggsave("total_districts.png", height = 4, width = 7)

# school_meals_average %>% 
#   filter(type == "percent_eligible_frpm_k12") %>% 
#   ggplot(aes(x = end_month, y = total_enroll)) +
#   geom_col()


```

```{r}

school_meals_average %>% 
  mutate(
    time_period = 
      if_else(
        end_month %in% c("June", "July"), "June and July", "August and beyond"
    ),
    time_period = factor(time_period, levels = c("June and July", "August and beyond"))
  ) %>% 
  group_by(time_period) %>% 
  summarize(
    total_frac_enroll = sum(total_frac_enroll),
    total_frac_frpm_eligible = sum(total_frac_frpm_eligible)
  ) %>% 
  mutate(percent_eligible_frpm_k12 = total_frac_frpm_eligible / total_frac_enroll) %>% 
  ggplot(aes(x = time_period, y = percent_eligible_frpm_k12)) +
  geom_col() +
  ggrepel::geom_text_repel(
    aes(label = scales::percent(percent_eligible_frpm_k12)),
    direction = "y",
    nudge_y = 0.01,
    size = 3.5
  ) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
  coord_cartesian(expand = TRUE) +
  labs(
    title = "Average percentage of children eligible for FRPM in districts closing meal sites",
    y = "Average percentage children eligible for Free-Reduced Meals",
    x = "Time Period",
    caption = "Source: CA Dept of Ed; BayAreaCommunity.org"
  )

ggsave("timeperiods.png", height = 4, width = 7)
```

